---
title: "Aztec DH Assignments"
author: "Ahmad ElBaba"
date: "2023-12-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing and loading necessary packages:
```{r}
install.packages("devtools")
devtools::install_github("statsbomb/SDMTools")
devtools::install_github("statsbomb/StatsBombR")
library('StatsBombR')
library(ggplot2)
library(SBpitch)
```

Assignment 1:  Use StatsBomb's free event data on the FA Women's Super League 2020/2021 Season. Assume you are Manchester City and you're going up against Chelsea. Build an optimal, data-driven attacking corner kick session that you'd implement in practice. Explain your thought process, data approach, algorithm performance.

Before starting the analysis, I brainstormed a plan on how I should proceed with the analysis. I've decided it to split it up into 2 part. 

1) Identifying patterns and player analysis:
Assessing areas of strength and weaknesses in Man City's corner kicks. Which types of corner kicks are more effective(inswinging vs outswinging, high cross vs low cross vs ground cross).
Assessing Chelsea's ability to defend corners(Goalkeeper positioning)
Which Man City players are causing the greatest threat to goal from corners(Scoring and crossing).

2) Designing the drill
Define objectives from the drills(e.g improve crosses/improve scoring)
Design the drills specific to Chelsea's defending

Loading our dataset, filtering our data for passes and corners:
```{r}
comp <- FreeCompetitions() %>% filter(competition_id==37 & season_name=="2020/2021")
comp 
matches <- FreeMatches(comp)
StatsBombData <- free_allevents(MatchesDF = matches, Parallel = T) 
StatsBombData = allclean(StatsBombData)
passes = StatsBombData %>%
  filter(type.name=="Pass" & is.na(pass.outcome.name))
cornerKicks = passes %>% filter(pass.type.name == "Corner") 
cornerKicks_mancity = cornerKicks %>%  filter(team.name == "Manchester City WFC")
```


Splitting the corners based on their height to see which is most effective:
```{r}
low_corners = cornerKicks_mancity %>% filter(pass.height.name == "Low Pass")
crossed_corners = cornerKicks_mancity %>% filter(pass.height.name == "High Pass")
ground_corners = cornerKicks_mancity %>%  filter(pass.height.name == "Ground Pass")
```

Creating field visuals to see where the corners were played:
Low Corners:
```{r}
create_Pitch() +
  geom_segment(data = low_corners, aes(x = location.x, y = location.y,
                                  xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Low corner kicks", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100) 
```

Ground Corners:
```{r}
create_Pitch() +
  geom_segment(data = ground_corners, aes(x = location.x, y = location.y,
                                       xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Ground corner kicks", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100) 
```

High Corners
```{r}
create_Pitch() +
  geom_segment(data = crossed_corners, aes(x = location.x, y = location.y,
                                               xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's High corner kicks", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100) 
```

Visualizing the percentage of corners by height:
```{r}
percentage <- cornerKicks_mancity %>% count(pass.height.name) %>% mutate(Percentage = n / sum(n) * 100)
ggplot(percentage, aes(x = "", y = Percentage, fill = pass.height.name)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(fill = "Corner Height") +geom_text(aes(label=Percentage), position = position_stack(vjust = 0.5))

```

High corners are the most common and low corners are the least. High corners are more direct and may lead to higher goal threat. Ground corners are often played short, but other event may come next which leads to a goal.

Determining the successes of corners by whether the goal was scored or not
goalfrom_corners means the corner directly resulted in a goal
successful_corners means the corner reached a teammate but wasn't scored
unsuccessful_corners means the corner didn't reach a teammate. Either cleared by opposition or went out of bounds
```{r}
goalfrom_corners = cornerKicks_mancity %>% filter(pass.shot_assist == TRUE) 
successful_corners = cornerKicks_mancity %>% filter(is.na(pass.shot_assist)) 
unsuccessful_corners = cornerKicks_mancity %>% filter(is.na(pass.recipient.name)) 
```

35 goals from corners
49 corners with recipient but no assist recorded
All corners taken by Man City reached a teammate

Creating pitch visuals for goals from corner and successful corners
```{r}
create_Pitch() +
  geom_segment(data = goalfrom_corners, aes(x = location.x, y = location.y,
                                           xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Goals from corners", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)

create_Pitch() +
  geom_segment(data = successful_corners, aes(x = location.x, y = location.y,
                                            xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Successful corners", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)
```

Just by visually analysing these graphics, we can see that corners in the box result in the higher success rate. Closer to goal results in a higher chance to score. 

Now let's check whether inswingers or outswingers are more effective. Let's start by plotting their frequencies in a bar chart
```{r}
ggplot(cornerKicks_mancity, aes(x = pass.technique.name)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Manchester City's corner techniques", subtitle = "WSL, 
2020-21", x = "Corner Technique", y = "Count")

```
Inswinging seems to be the most used, followed by outswinging. Inswinging corners are closer to goal, which can increase the chance of scoring. However, it also means they're close to the goalkeeper, making it easier for the keeper to catch the goal. Let's see the area the corner each corner gets played into and their success rate in terms of percentage in a pie chart.

Inswinging:
```{r}
inswingingCorners = cornerKicks_mancity %>% filter(pass.technique.name == "Inswinging")
create_Pitch() +
  geom_segment(data = inswingingCorners, aes(x = location.x, y = location.y,
                                           xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Goals from corners", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)

percentage <- inswingingCorners %>% count(pass.shot_assist) %>% mutate(Percentage = n / sum(n) * 100)
ggplot(percentage, aes(x = "", y = Percentage, fill = pass.shot_assist)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title =  "Inswinging corners success", fill = "Goal from corner") +geom_text(aes(label=Percentage), position = position_stack(vjust = 0.5))
```

Outswinging:
```{r}
outswingingCorners = cornerKicks_mancity %>% filter(pass.technique.name == "Outswinging")
create_Pitch() +
  geom_segment(data = outswingingCorners, aes(x = location.x, y = location.y,
                                           xend = pass.end_location.x, yend = pass.end_location.y),
               lineend = "round", size = 0.5, colour = "#000000", arrow =
                 arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) +
  labs(title = "Manchester City Women's Goals from corners", subtitle = "WSL,
2020-21") + 
  scale_y_reverse() +
  coord_fixed(ratio = 105/100)

percentage <- outswingingCorners %>% count(pass.shot_assist) %>% mutate(Percentage = n / sum(n) * 100)
ggplot(percentage, aes(x = "", y = Percentage, fill = pass.shot_assist)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title =  "Outswinging corners success", fill = "Goal from corner") +geom_text(aes(label=Percentage), position = position_stack(vjust = 0.5))
```
Outswinging corners seem to be leading to more goals than inswinging. From the pitch visual, we can see that outswinging corners are usually further away from the keeper's 6 yard box. Let's see Chelsea's keeper's positioning and their ariel ability. 

```{r}
chelseaKeeper = StatsBombData %>% filter(team.name == "Chelsea FCW" & type.name == "Goal Keeper" & play_pattern.name == "From Corner" & goalkeeper.outcome.name != "Shot Faced")
ggplot(chelseaKeeper, aes(x = goalkeeper.type.name)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Chelsea Goalkeeprs ariel ability", subtitle = "WSL, 
2020-21", x = "Corner Technique", y = "Count")

```


```{r}
library(tidyverse)
heatmap = chelseaKeeper %>%
                    mutate(location.x = ifelse(location.x>120, 120, location.x),
                           location.y = ifelse(location.y>80, 80, location.y),
                           location.x = ifelse(location.x<0, 0, location.x),
                           location.y = ifelse(location.y<0, 0, location.y)) 
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE)
heatmap = heatmap%>%
  group_by(goalkeeper.type.name) %>%
  mutate(total_DA = n()) %>%
  group_by(goalkeeper.type.name, xbin, ybin) %>%
  summarise(total_DA = max(total_DA),
            bin_DA = n(),
            bin_pct = bin_DA/total_DA,
            location.x = median(location.x),
            location.y = median(location.y)) %>%
  group_by(xbin, ybin)

library(grid)
keepermovement_colors <- c("#dc2429","#ef9195", "#f9e0e2", "#c0c7cd", "#798590",
"#1d3048","#020c16") 
ggplot(data= heatmap, aes(x = location.x, y = location.y)) +
  geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + 
  annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
  annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
  annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
  annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
  theme(rect = element_blank(),
        line = element_blank()) +
  annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
  annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
  annotate("path", colour = "white", size = 0.6,
           x=60+10*cos(seq(0,2*pi,length.out=2000)),
           y=40+10*sin(seq(0,2*pi,length.out=2000)))+
  annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
  annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
           y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
  annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
           y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
  theme(axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption=element_text(size=13,family="sans", hjust=0.5, vjust=0.5),
        plot.subtitle = element_text(size = 18, family="sans", hjust = 0.5),
        axis.text.y=element_blank(),
        legend.title = element_blank(),
        legend.text=element_text(size=22,family="sans"),
        legend.key.size = unit(1.5, "cm"),
        plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 32.5,
                                  family="sans", colour = "black", hjust = 0.5),
        legend.direction = "vertical",
        axis.ticks=element_blank(),
        plot.background = element_rect(fill = "white"),
        strip.text.x = element_text(size=13,family="sans")) + 
  scale_y_reverse() + 
  scale_fill_gradientn(colours = keepermovement_colors, trans = "reverse", labels =
                         scales::percent_format(accuracy = 1), limits = c(0.03, -0.03)) + 
  coord_fixed(ratio = 95/100) +
  annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
                                                 length=unit(1.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
                    xmin=25, xmax = 95, ymin = -83, ymax = -83) +
  facet_wrap(~goalkeeper.type.name)+ 
  guides(fill = guide_legend(reverse = TRUE))

```

From corners directly, we can see that Chelsea's keepers are able to deal with them without conceding. The bar plot shows that their keepers are able to punch most of the corners away, while also saving or collecting some. From this season, they conceded 2 goals directly from corners. Therefore, outswinging corners seem to be the more effective corner technique, further away from the keeper. 

Let's see which Manchester City players cause the biggest aerial threat in front of goal. 
```{r}
corner_shots = StatsBombData %>% filter(play_pattern.name == "From Corner" & type.name == "Shot" & team.name == "Manchester City WFC") %>% count(player.name) %>% arrange(desc(n))

print(head(corner_shots))

```


```{r}
corner_goals = StatsBombData %>% filter(play_pattern.name == "From Corner" & type.name == "Shot" & team.name == "Manchester City WFC") %>% filter(shot.outcome.name == "Goal") %>% count(player.name) %>% arrange(desc(n))

print(head(corner_goals))
```
Lucy Bronze, Ellen White and Samantha Mewis seems to be among the most threatening players from corners. Hence, there should be a focus on having them in the box for corner kicks.

Let's take a look at who City's corner kick takers should be

```{r}
cornerkick_takers.right = cornerKicks_mancity %>% filter(pass.body_part.name == "Right Foot") %>%  count(player.name) %>% arrange(desc(n))

print(head(cornerkick_takers.right))
```


```{r}
cornerkick_takers.left = cornerKicks_mancity %>% filter(pass.body_part.name == "Left Foot") %>%  count(player.name) %>% arrange(desc(n))

print(head(cornerkick_takers.left))
```

Chloe Kelly to take left corners and Alex Greenwood to take right corners since we chose to make use of outswinging corners.

```{r}
after_groundpass = StatsBombData %>% filter(play_pattern.name == "From Corner" & team.name == "Manchester City WFC") %>% count(type.name) %>% arrange(desc(n))

print(after_groundpass)
```

```{r}
shots_groundpass = StatsBombData %>% filter(play_pattern.name == "From Corner" & team.name == "Manchester City WFC" & type.name == "Shot") %>% count(shot.outcome.name) %>% arrange(desc(n))

print(shots_groundpass)
```
```{r}
percentage <- shots_groundpass %>% count(shot.outcome.name) %>% mutate(Percentage = n / sum(n) * 100) 
ggplot(percentage, aes(x = "", y = Percentage, fill = shot.outcome.name)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(title =  "Shots after ground corners", fill = "Shot outcomes") +geom_text(aes(label=Percentage), position = position_stack(vjust = 0.5))
```

Ground passes for Manchester City have a 14% success rate. High corners had the highest success rate, therefore, high corners should be favored. 

After analyzing the patterns and player analysis, we can start defining the objectives of the drill. Seeing that most corner kicks were accurate and reaching the players, but shot accuracy was lower (some players having 10-12 shots but scoring a max of 3), there could be an emphasis of improving finishing in Man City's training with players such as Lucy Bronze, Ellen White and Samantha Mewis being the focus and Chloe Kelly taking left corners and Alex Greenwood taking the right.

With outswinging high corners being the most successful, there should also be an aim to practice those as well to make sure they're perfected. An area of improvement should also be ground corners which only resulted in 14% of shots scored. While outswinging corners can be effective in this case, using them all the time would make it too predictable for Chelsea's defense. Since Chelsea's keepers seem capable of dealing with corners, inswinging corners may not be as helpful. Therefore, being able to rely on short ground passes would be beneficial for City. Designing a specific drill for ground corners would help cause a bigger goal threat. 

Key takeaways:
-High corners and Outswinging corners are a strength 
-Improve ground corners
-Chloe Kelly and Alex Greenwood take corners
-Lucy Bronze, Ellen White Samantho Mewis Aerial Goal threats
-Improve finishing from corners

Assignment 2: Is event data good for evaluating defending, yes or no? Explain why.

Yes, event data can be good for evaluating defending. It would allow teams to see the strengths and weaknesses of the defense. For example, the offensive team can take advantage of weaknesses in the defense. Let's take the event data given to us from the 2020-21 WSL season. We can build a heatmap of defensive activities of each team in the WSL.

```{r}

heatmap = StatsBombData %>%mutate(location.x = ifelse(location.x>120, 120, location.x),
                           location.y = ifelse(location.y>80, 80, location.y),
                           location.x = ifelse(location.x<0, 0, location.x),
                           location.y = ifelse(location.y<0, 0, location.y)) #1
heatmap$xbin <- cut(heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
heatmap$ybin <- cut(heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE)
heatmap = heatmap%>%
  filter(type.name=="Pressure" | duel.type.name=="Tackle" |
           type.name=="Foul Committed" | type.name=="Interception" |
           type.name=="Block" ) %>%
  group_by(team.name) %>%
  mutate(total_DA = n()) %>%
  group_by(team.name, xbin, ybin) %>%
  summarise(total_DA = max(total_DA),
            bin_DA = n(),
            bin_pct = bin_DA/total_DA,
            location.x = median(location.x),
            location.y = median(location.y)) %>%
  group_by(xbin, ybin) %>%
  mutate(league_ave = mean(bin_pct)) %>%
  group_by(team.name, xbin, ybin) %>%
  mutate(diff_vs_ave = bin_pct - league_ave)

defensiveactivitycolors <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51",
                             "#e35256", "#e76266", "#e9777b", "#ec8589", "#ec898d", "#ef9195",
                             "#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2",
                             "#f7e1e3", "#f5e2e4", "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd",
                             "#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98", "#798590",
                             "#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048",
                             "#11263e", "#11273e", "#0d233a", "#020c16") 
ggplot(data= heatmap, aes(x = location.x, y = location.y, fill = diff_vs_ave, group =diff_vs_ave)) +
  geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
  annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
  annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
  annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
  annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
  annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
  annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
  theme(rect = element_blank(),
        line = element_blank()) +
  annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
  annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
  annotate("path", colour = "white", size = 0.6,
           x=60+10*cos(seq(0,2*pi,length.out=2000)),
           y=40+10*sin(seq(0,2*pi,length.out=2000)))+
  annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
  annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
           y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
  annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
           y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
  theme(axis.text.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.caption=element_text(size=13,family="Source Sans Pro", hjust=0.5, vjust=0.5),
        plot.subtitle = element_text(size = 18, family="Source Sans Pro", hjust = 0.5),
        axis.text.y=element_blank(),
        legend.title = element_blank(),
        legend.text=element_text(size=22,family="Source Sans Pro"),
        legend.key.size = unit(1.5, "cm"),
        plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 32.5,
                                  family="Source Sans Pro", colour = "black", hjust = 0.5),
        legend.direction = "vertical",
        axis.ticks=element_blank(),
        plot.background = element_rect(fill = "white"),
        strip.text.x = element_text(size=13,family="Source Sans Pro")) + #4
  scale_y_reverse() + #5
  scale_fill_gradientn(colours = defensiveactivitycolors, trans = "reverse", labels =
                         scales::percent_format(accuracy = 1), limits = c(0.03, -0.03)) + #6
  labs(title = "Where Do Teams Defend vs League Average?", subtitle = "FA Women's Super
League, 2020/21") + #7
  coord_fixed(ratio = 95/100) +
  annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
                                                 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
                    xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
  facet_wrap(~team.name)+ #10
  guides(fill = guide_legend(reverse = TRUE))


```

From this heatmap, we can see that teams such as Bristol City defend more within their defensive third, whereas Manchester City press more in the opposition third. Teams like Arsenal tend to defend more on the right side, whereas Brighton defend more on the left. Teams can see such heatmaps and tailor their offensive style based on the opposing teams defense. 

We can also use event data to see how aggressive each team plays. From events such as fouls, yellow cards and red cards, we can analyze the aggressiveness of each player and each team. Let's look at those stats for the 2020-21 season.

```{r}
fouls = StatsBombData %>% filter(type.name == "Foul Committed" & foul_committed.type.name != "Handball")

ggplot(fouls, aes(x = team.name)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Total Number of Fouls", 
       subtitle = "WSL, 2020-21", 
       x = "Team Name", 
       y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))  

```

```{r}

ggplot(fouls, aes(x = foul_committed.card.name)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Total Number of Fouls", 
       subtitle = "WSL, 2020-21", 
       x = "Team Name", 
       y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
    facet_wrap(~team.name)
```

```{r}
fouls_player = fouls %>%  filter(foul_committed.card.name == "Yellow Card") %>%  count(player.name) %>% arrange(desc(n))
 
print(fouls_player)
```

From this event data, we can see which teams and players have the most aggressive play style in the league. This can allow coaches to change their play style depending on the team they come up against. It can also help coaches know which players are more prone to being booked or sent off.

Event data can also be helpful for comparing players when scouting. For example, let's assume we are one of the teams in the WSL and we have defensive problems in terms of aerial duels. Event data helps us compare our players with other players in the league with the aims of improving our defense. Let's take an example from our event data. We can look at which team has lost the most Aerial duels in the season

```{r}
aerial_duels = StatsBombData %>% filter(duel.type.name == "Aerial Lost") %>% count(team.name) %>% arrange(desc(n))

print(aerial_duels)
```

Let's assume we are West Ham and we want to improve our problem with Aerial Duels. Let's look at data from the team with the least duels lost.

```{r}
aerial_duels_least = StatsBombData %>% filter(duel.type.name == "Aerial Lost") %>% count(team.name) %>% arrange(n)

print(aerial_duels_least)
```

```{r}
aerial_duels_arsenal = StatsBombData %>% filter(duel.type.name == "Aerial Lost") %>%  filter(team.name == "Arsenal WFC")

ggplot(aerial_duels_arsenal, aes(x = duel.type.name)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Total Number of Aerial Duels Lost", 
       subtitle = "Arsenal WFC, WSL, 2020-21", 
       x = "Player Name", 
       y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
    facet_wrap(~player.name)
```

From this Arsenal team, we can see that Lisa Evans, Stepanie-Elise Catley and Viktoria Schnaderbeck could be possible players that can solve our problem with Aerial duels. Of course, there are other factors to consider such as budget and other attributes our players might have that the players we are targetting won't. However, this is just looking at it from the perspective of aerial duels. 

These are just a few ways event data could be useful for evaluating defending. 
